Learning Constraints and Optimization Criteria

### 2015-10-23 Paper.tex
Import own background change
Import Anton changes approach

\section{Approach}
In this research, both constraint learning and learning optimization criteria were implemented.
A two-step approach is used to learn optimization criteria.
First, clauses learning is used to find soft constraints in the given examples.
Then, user-provided preference information is used to learn weights for these constraints.

@@

\subsection{Input}
The input consists of global definitions that specify the types and predicates that are used to describe examples and a set of examples \sym{E}.
Typing information relates directly to the problem domain and is usually easy to provide.
Including types also improves the accuracy and efficiency of the learned clauses.

Examples are interpretations, they specify the constants in their domain and then all predicates that are true.
Predicates can also be generated by background knowledge, in which case they are not included in the example.

When learning optimization criteria, preference information is also provided by the user.
Each preference is of the form $e_1 \succ e_2 \succ ... \succ e_n$ with $e_i \in \sym{E}$.
Preference look at the relative position of a few examples at a time and are typically easier to provide than absolute scores for all examples.

@@

\subsection{Clausal Discovery}
The clause learning system is based on the clausal discovery algorithm (alg.~\ref{alg:cd}).
Note the double use of the $\models$ operator.
On the one hand, it is used in $e \models c$ to denote that clause~$c$ is true in the example (interpretation)~$e$ and $c$ is said to cover $e$.
This test is used to compute if a clause covers enough examples.
On the other hand, in $\sym{T} \models c$ it signifies that \sym{T} entails $c$ (i.e. for every interpretation it holds that if \sym{T} is true, $c$ is also true).
The algorithm uses this test to remove redundant clauses.
For both cases the IDP knowledge base system \cite{de2013prototype,wittocx2008idp} is used to compute $\models$.
The algorithm can find soft constraints ($t < \# \sym{E}$) or hard constraints ($t = \# \sym{E}$).

Starting with the empty clause ($\square$), clauses are refined until they cover enough examples.
If a clause covers enough examples it is added to result set, provided it is not entailed by clauses in the result set.
In the refinement step, a clause $c = \{l_1, l_2, ..., l_k\}$ is extended by adding a new literal $l$.
The new clause, $c' = c \cup \{l\}$ is more than general than $c$ ($c' \models c$) and potentially covers more examples.
By adding just one literal the refinement operator $\rho$ generates a set of clauses that are minimally more general.
In order to do this efficiently, a list of atoms is pre-calculated in advance, given a maximal amount of variables that can occur within a clause.
Within the body and head of a clause, atoms may only be added in a specific order to avoid generating redundant clauses.
The implementation natively supports symmetric predicates for which the order of the terms do not matter.
Object Identity is used to specify that variables with different names must denote different objects.
Additionally there are two syntactic restrictions.
Clauses must be connected (i.e. new atoms must always contain a variable that has already occurred) and range-restricted (i.e. no new variables may be introduced in the head of the clause).

To avoid extensive coverage ($e \models c$) and entailment ($\sym{T} \models c$) calculations, two additional tests are used.
The subset-test rejects clauses that are supersets of a clause that covers the same or all examples and has been accepted earlier.
Because clauses can be formulated in multiple ways, a representative-test is used to remove clauses that are not in canonical form.

@@

\subsection{Optimization}
The first step in finding weighted soft constraints is to identify soft constraints by using the constraint learning system with a (low) threshold.
Examples are characterized by the soft constraints (clauses) which they satisfy.
Therefore, every example~$e$ can be translated to a vector of boolean features by introducing a feature~$f_i$ for every clause~$c_i$, with $f_i = \mathbf{1}_{c_i \text{ covers } e}$.
Existing software can be used to learn a linear scoring function $\sum_i w_i \cdot f_i$ over these features based on the given rankings.
For an unseen example, the feature vector is computed by calculating what clauses cover the new example.
The scoring function can then be used to calculate a score for that example.

\begin{example} "Rewrite"
  Consider examples~$e_1, e_2, e_3$, rankings $e_1 \succ e_2, e_2 \succ e_3$ and clauses $c_1, c_2, c_3$.
  If the clauses cover examples according to table~\ref{tbl:cover_examples}, then, for example, the function $(1 \cdot c_1) + (0\cdot c_2) + (-2\cdot c_3)$ perfectly models the given rankings.
  A new example that is covered by~$c_1$ and~$c_2$, would be assigned a score of $1$, according to this function.
  This score has no value in the absolute sense, it can only be used to compare it to other examples ranked by the same function.
  In this example clause~$c_1$ represents a desirable property, clause~$c_2$ is ignored because it does not influence the ranking and clause~$c_3$ represents an undesirable property.

@@

Using $t$ and $\mathit{cost}$, a function $\mathit{actual}(i)$ is then defined in IDP as $\mathbf{1}_{\lnot t(i)} \cdot \mathit{cost}(i)$.
This function is used in the term $\sum_i actual(i)$ to be minimized, which will allow IDP to search for an optimal solution.

>>>>

\section{Approach}
% In this research, both constraint learning and learning optimization criteria were implemented.
In this work, we use a two-step approach to learn optimization criteria.
First, clause learning is used to find soft constraints in the given examples.
Then, user-provided preference information is used to learn weights for these constraints.

@@

\subsection{Input}
The input consists of global definitions that specify the types and predicates that are used to describe examples and a set of examples \sym{E}.
\anton{Explain: what is type information?}
Type information relates directly to the problem domain and is usually easy to provide.
Including types also improves the accuracy and efficiency of the learned clauses.

Examples are interpretations, they specify the constants in their domain and all predicates that are true.
Predicates can also be generated by background knowledge, in which case they are not included in the example.

When learning optimization criteria, preference information is also provided by the user.
Each preference is of the form $e_1 \succ e_2 \succ ... \succ e_n$ with $e_i \in \sym{E}$.
Preferences describe the relative position of a few examples at a time and are typically easier to provide than absolute scores for all examples.

@@

\subsection{Clausal Discovery}
The clause learning system is based on the clausal discovery algorithm (alg.~\ref{alg:cd}).
Note the double use of the $\models$ operator. \anton{This paragraph seems strange.}
On the one hand, it is used in $e \models c$ to denote that clause~$c$ is true in the example (interpretation)~$e$ and $c$ is said to cover $e$.
This test is used to compute if a clause covers enough examples.
On the other hand, in $\sym{T} \models c$ it signifies that \sym{T} entails $c$ (i.e. for every interpretation it holds that if \sym{T} is true, $c$ is also true).
The algorithm uses this test to remove redundant clauses.
For both cases the IDP knowledge base system \cite{de2013prototype,wittocx2008idp} is used to compute $\models$.
The algorithm can find soft constraints ($t < \# \sym{E}$) or hard constraints ($t = \# \sym{E}$).

Starting with the empty clause ($\square$), clauses are refined until they cover enough examples.
If a clause covers enough examples it is added to the result set, provided it is not entailed by other clauses already in the result set.
In the refinement step, a clause $c = \{l_1, l_2, ..., l_k\}$ is extended by adding a new literal $l$.
The new clause, $c' = c \cup \{l\}$ is more general than $c$ ($c' \models c$) and potentially covers more examples.
By adding just one literal the refinement operator $\rho$ generates the set of clauses that are minimally more general.
In order to do this efficiently, a list of atoms is pre-calculated, given a maximal amount of variables that can occur within a clause.
Within the body and head of a clause, atoms may only be added in a specific order to avoid generating redundant clauses.
The implementation natively supports symmetric predicates for which the order of the terms do not matter.
Object Identity is used to specify that variables with different names must denote different objects.
Additionally there are two syntactic restrictions.
Clauses must be connected (i.e. new atoms must always contain a variable that has already occurred) and range-restricted (i.e. no new variables may be introduced in the head of the clause).
\anton{This is a vague list of unconnected statements.}

To avoid extensive coverage ($e \models c$) and entailment ($\sym{T} \models c$) calculations, two additional tests are used.
The subset-test rejects clauses that are supersets of a clause that covers the same or all examples and has been accepted earlier.
Because clauses can be formulated in multiple ways, a representative-test is used to remove clauses that are not in canonical form.
\anton{Not clear, needs more explanation, or should be removed.}

@@

\subsection{Optimization}
The first step in finding weighted soft constraints is to identify soft constraints by using the constraint learning system with a (low) threshold.
Examples are characterized by the soft constraints (clauses) which they satisfy.
Therefore, every example~$e$ can be translated to a vector of boolean features by introducing a feature~$f_i$ for every clause~$c_i$, with $f_i = \mathbf{1}_{c_i \text{ covers } e}$. \anton{What does this mean?}
Existing software can be used to learn a linear scoring function $\sum_i w_i \cdot f_i$ over these features based on the given rankings.
For an unseen example, the feature vector is computed by calculating what clauses cover the new example.
The scoring function can then be used to calculate a score for that example.

\begin{example} "Rewrite"
  Consider examples~$e_1, e_2, e_3$, rankings $e_1 > e_2, e_2 > e_3$ and clauses $c_1, c_2, c_3$.
  If the clauses cover examples according to table~\ref{tbl:cover_examples}, then, for example, the function $(1 \cdot c_1) + (0\cdot c_2) + (-2\cdot c_3)$ perfectly models the given rankings.
  A new example that is covered by~$c_1$ and~$c_2$, would be assigned a score of $1$, according to this function.
  This score has no value in the absolute sense, it can only be used to compare it to other examples ranked by the same function.
  In this example clause~$c_1$ represents a desirable property, clause~$c_2$ is ignored because it does not influence the ranking and clause~$c_3$ represents an undesirable property.

@@

Using $t$ and $\mathit{cost}$, a function $\mathit{actual}(i)$ is then defined in IDP as $\mathbf{1}_{\lnot t(i)} \cdot \mathit{cost}(i)$.
This function is used in the optimization criterion $\sum_i actual(i)$ which should be minimized. % , which will allow IDP to search for an optimal solution.
